A Dependently Typed Multi-Stage Calculus \ifnotanonymous\thanks{Supported by organization x.}\fi

### 2019-06-11 main.tex
Revise until Contexts.

\begin{align*}
	\textrm{Type variables}  &  &                          & X,Y,Z                                                                             \\
	\textrm{Variables}       &  &                          & x,y,z                                                                             \\
	\textrm{Stage variables} &  &                          & \alpha,\beta,\gamma                                                               \\
	\textrm{Stage}           &  &                          & A,B,C                                                                             \\
	\textrm{Kinds}           &  & K,J,I,H,G                & ::= * \mid \Pi x:\tau.K                                                           \\
	\textrm{Types}           &  & \tau,\sigma,\rho,\pi,\xi & ::= X \mid \Pi x:\tau.\tau \mid \tau\ M \mid \TW_{\alpha} \tau \mid \F\alpha.\tau \\
	\textrm{Constants}       &  &                          & c                                                                                 \\
	\textrm{Terms}           &  & M,N,L,O,P                & ::= c \mid x \mid \lambda x:\tau.M\ \mid M\ M \mid \TB_\alpha M                   \\
	                         &  &                          & \ \ \ \ \mid \TBL_\alpha M \mid \Lambda\alpha.M \mid M\ \epsilon \mid \%_\alpha M         \\  
	\textrm{Signature}       &  & \Sigma                   & ::= \emptyset \mid X::K \mid c:\tau                                               \\
	\textrm{Contexts}        &  & \Gamma                   & ::= \emptyset \mid  \Gamma,x:\tau @A\ (x\not\in\textrm{Domain}(\G))               \\
\end{align*}

@@

A type is a type variable which is declared in the signature, a dependent function type,
an application of a term to a dependent function type, a code type, or an $\alpha$-closed type.
A function type in ordinary type systems is omitted 
because it is a special case of a dependent type $\Pi x:\tau$, which has no free occurrence of $x$  in $\tau$.
An application of a term to a dependent function type is, for example, $\text{Vector}\ 10$
if type variable $\text{Vector}$ has the kind $\Pi x:\I.*$.
A code type $\TW_\alpha \tau$ denotes a code fragment of a term of type $\tau$.
An $\alpha$-closed type corresponds to a runnable code fragment.
Type constructors tie strongly in the order of 
an application of a term to a dependent function type $\tau\ M$, a dependent function type $\Pi x:\tau.\tau$, 
a code type $\TW_{\alpha} \tau$, an $\alpha$-closed type $\F\alpha.\tau$
Therefore, $\F\alpha.\TW_{\alpha}.\Pi x:\I.\text{Vector}\ 5$ is interpreted as
$\F\alpha.(\TW_{\alpha}.(\Pi x:\I.(\text{Vector}\ 5)))$.

\red{ 
$c$ ranges over constants, 
}
In addition to ordinary terms in Simply Typed Lambda Calculus, there are five more forms.
$\TB_\alpha M$ represents a code fragment, and $\TB_\alpha M$ represents unquote.
$\Lambda\alpha.M$ is a stage variable abstraction.
$M\ \epsilon$ is an application of $\epsilon$ to a stage variable.
$\%_\alpha M$ is a primitive operator for cross-stage persistence.

Signatures are sequences of pairs of a constant and its type or a type variable and its kind.
Because we adopted a \LLF-like system, constants and type variables for base types are given in the signature $\Sigma$.
For example, when we use integers in \LMD, $\Sigma = \I :: *, 1:\I, 2:\I, \cdots$.

% Contexts
Contexts are sequences of triplexes of a variable, its type, and its stage.
In order to restrict contexts to only well-formed ones, there is a condition of $x\notin\FV(\G)$.
Because of this condition, there is no type with free variables in contexts.

% 演算子の結合順序
Connection degree of $\TB_\alpha, \TBL_\alpha, \%_\alpha$ is stronger than the two forms of applications
and applications are left-associative
and two abstractions extends as far to the right as possible.

% 自由変数
As usual, the variable $x$ is bound in $\lambda x:\tau.M$
and the transition variable $\alpha$ is bound in $\Lambda \alpha.M$.
We identify $\alpha$-convertible terms and assume the names of bound variables are pairwise distinct.
We write $\FV(M)$ and $\FTV(M)$ for the set of free variables and the set of free stage variables in $M$, respectively.
We omit their definitions.

@@

>>>>

\begin{align*}
	\textrm{Type variables}  &  &                          & X,Y,Z                                                                             \\
	\textrm{Variables}       &  &                          & x,y,z                                                                             \\
	\textrm{Stage variables} &  &                          & \alpha,\beta,\gamma                                                               \\
	\textrm{Stage}           &  &                          & A,B,C                                                                             \\
	\textrm{Kinds}           &  & K,J,I,H,G                & ::= * \mid \Pi x:\tau.K                                                           \\
	\textrm{Types}           &  & \tau,\sigma,\rho,\pi,\xi & ::= X \mid \Pi x:\tau.\tau \mid \tau\ M \mid \TW_{\alpha} \tau \mid \F\alpha.\tau \\
	\textrm{Constants}       &  &                          & c                                                                                 \\
	\textrm{Terms}           &  & M,N,L,O,P                & ::= c \mid x \mid \lambda x:\tau.M\ \mid M\ M \mid \TB_\alpha M                   \\
	                         &  &                          & \ \ \ \ \mid \TBL_\alpha M \mid \Lambda\alpha.M \mid M\ \epsilon \mid \%_\alpha M         \\  
	\textrm{Signature}       &  & \Sigma                   & ::= \emptyset \mid X::K \mid c:\tau                                               \\
	\textrm{Contexts}        &  & \Gamma                   & ::= \emptyset \mid  \Gamma,x:\tau @A\ (x\not\in\textrm{Domain}(\G)\ \text{and}\ x\not\in\textrm{FV}(\tau))               \\
\end{align*}

@@

A type is a type variable which is declared in the signature, a dependent function type,
an application of a term to a dependent function type, a code type, or an $\alpha$-closed type.
A function type in ordinary type systems is omitted 
because it is a special case of a dependent type $\Pi x:\tau$, which has no free occurrence of $x$  in $\tau$.
An application of a term to a dependent function type is, for example, $\text{Vector}\ 10$
if type variable $\text{Vector}$ has the kind $\Pi x:\I.*$.
A code type $\TW_\alpha \tau$ denotes a code fragment of a term of type $\tau$.
An $\alpha$-closed type corresponds to a runnable code fragment.
Type constructors tie strongly in the order of 
an application of a term to a dependent function type $\tau\ M$,
a code type $\TW_{\alpha} \tau$,
a dependent function type $\Pi x:\tau.\tau$, 
an $\alpha$-closed type $\F\alpha.\tau$.
Therefore, $\F\alpha.\TW_{\alpha}.\Pi x:\I.\text{Vector}\ 5$ is interpreted as
$\F\alpha.(\TW_{\alpha}.(\Pi x:\I.(\text{Vector}\ 5)))$.

% Terms

Terms include a constant $c$ which is declared in signature $\Sigma$.
In addition to ordinary terms in Simply Typed Lambda Calculus, there are five more forms.
$\TB_\alpha M$ represents a code fragment, and $\TB_\alpha M$ represents escape.
$\Lambda\alpha.M$ is a stage variable abstraction.
$M\ \epsilon$ is an application of $\epsilon$ to a stage variable.
$\%_\alpha M$ is a primitive operator for cross-stage persistence.
$\TB_\alpha, \TBL_\alpha, \%_\alpha$ tie stronger than the two forms of applications
and applications are left-associative
and two abstractions extends as far to the right as possible.
For example, $\F\alpha.\lambda x:\I.\TB_\alpha x\ y$ means $\F\alpha.(\lambda x:\I.(\TB_\alpha x)\ y)$.

% Signature

Signatures are sequences of pairs of a type variable and its kind or a constant and its type.
Because we adopted a \LLF-like system, constants and type variables for base types are given in the signature $\Sigma$.
For example, when we use integers in \LMD, $\Sigma = \I :: *, 1:\I, 2:\I, \cdots$.

% 自由変数
As usual, the variable $x$ is bound in $\lambda x:\tau.M$
and the transition variable $\alpha$ is bound in $\Lambda \alpha.M$.
We identify $\alpha$-convertible terms and assume the names of bound variables are pairwise distinct.

% Contexts
Contexts are sequences of triples of a variable, its type, and its stage.
We write $\FV(M)$ and $\FTV(M)$ for the set of free variables and the set of free stage variables in $M$, respectively.
$\textrm{Domain}$ is a function from a content to a set of variables defined in the context.
The side condition of $x\not\in\textrm{Domain}(\G)$ prevent multiple appearances of the same variable and
$x\not\in\textrm{FV}(\tau)$ is needed in the proof of Substitution Lemmas.

@@

\section{Appendix}
